---
title: "Project"
output: pdf_document
date: "2022-11-28"
---

```{r}
library(tidyverse)
library(ggplot2)
library(ggcorrplot)
library(car)
library(glmnet)
library(modelr)

```

## Description of data

```{r}
bd_df <- readxl::read_excel("data/body_density_data.xlsx")
dim(bd_df)
bd_df <- bd_df%>%
  select(bodyfat_brozek, age:wrist)%>%
  filter(bodyfat_brozek != 0)
```

We first looked into the content of the dataset. It has 252 observations across 
17 columns. The outcome we selected is bodyfat_brozek (body fat calculated by 
brozek), the other key variables are age, height, weight and circumference of 
body part like neck, chest, abdomen, hip, thigh, knee, ankle, bicep, forearm and wrist.

## plots
And we use plots to check the distribution of our variables and relationship 
between our outcome and each of the variables.

```{r}
lapply(bd_df, hist)
```


```{r}
par(mfrow = c(2,2))
lapply(colnames(bd_df)[2:length(colnames(bd_df))], function(nm){
  ggplot(bd_df) +
    geom_point(aes_string(y =colnames(bd_df)[1],
                          nm))
})
```

From the plots, it seems that all data are symmetrically distributed. We also saw that there is no obvious relation ship between height and body fat. 

```{r}
set.seed(1)
sub<-sample(nrow(bd_df),round(nrow(bd_df)*0.8))
data_train<-bd_df[sub,]
data_test<-bd_df[-sub,]
X.test = as.matrix(data_test[,-1]) 
Y.test = as.matrix(data_test[,1])
X.train = as.matrix(data_train[,-1]) 
Y.train = as.matrix(data_train[,1])
```

## Use stepwise methods
```{r}
full_fit = lm(bodyfat_brozek ~ age + height +weight + neck + chest + abdomen + hip + thigh + knee + ankle + bicep +forearm + wrist, data = data_train)
Step_model_1 = step(full_fit, direction = "backward", trace = 0)
summary(Step_model_1)
vif(Step_model_1)
plot(Step_model_1)
RMSE_Step <- rmse(Step_model_1,data = data_test)
RMSE_Step


```
Although circumference of hip is included in the stepwise model, it is not significant. We then calculated VIF in our stepwise model and found that the VIF of hip is >10. So we doubted that whether it can be included in a model.

```{r}
model2 <- lm(bodyfat_brozek ~ age  +weight + neck + abdomen  + thigh  +forearm + wrist, data = data_train)
anova(Step_model_1,model2)
summary(model2)
RMSE_2 <- rmse(model2,data = data_test)
RMSE_2
```

We performed ANOVA on these two models, p-value is greater than 0.05. The ANOVA test indicates that the model not include hip circumference may be better. This model also has a lower BIC. However, it's AIC is larger than the stepwise model and adj R-squared is also a little bit smaller than the stepwise model. Therefore, we can't decided our final model yet.

```{r}
mat = as.matrix(data_train)
leaps::leaps(x = mat[,-1], y = mat[,1], nbest = 1, method = "Cp")
leaps::leaps(x = mat[,-1], y = mat[,1], nbest = 1, method = "adjr2")
```

If we choose predictors based on Cp, The Cp is the smallest when the predictors are age, neck,chest, abdomen,hip, thigh, forearm and wrist. If we choose predictors based on adjust R-squared, the predictors are the same.

fitted model based on above Criterion
```{r}
model3 <- lm(bodyfat_brozek ~ age +  neck + chest + abdomen + hip + thigh  +forearm + wrist, data = data_train)
summary(model3)
RMSE_3 <- rmse(model3,data = data_test)
RMSE_3
```



## LASSO
We also want to use LASSO to generate a reasonable model. First, we want to 
choose the best lambda. So we conducted a cross validation. The cross 
validation's result shows that the best lambda is 0.04,which is small. 
This indicates that we may include many variables in our model if we using LASSO. 
So we fit a model with LASSO Regression.  
```{r}
set.seed(123)
lambda_seq <- 10^seq(-3, 0, by = .1)
cv_object <- cv.glmnet(as.matrix(data_train[,-1]), data_train$bodyfat_brozek, lambda = lambda_seq, 
nfolds = 5)
cv_object

tibble(lambda = cv_object$lambda,
mean_cv_error = cv_object$cvm) %>%
ggplot(aes(x = lambda, y = 
mean_cv_error)) +
geom_point()
cv_object$lambda.min
```

```{r}
fit_bestcv <- glmnet(as.matrix(data_train[,-1]), data_train$bodyfat_brozek, lambda = cv_object$lambda.min)
best_lambda = cv_object$lambda.min
coef(fit_bestcv)
```

```{r}
lasso_predict_train = predict(fit_bestcv, s = best_lambda, newx = X.train)
lasso_predict_test = predict(fit_bestcv, s = best_lambda, newx = X.test)
RMSE_LASSO = sqrt(mean((Y.test-lasso_predict_test)^2))
SSE_LASSO = sum((Y.train - lasso_predict_train )^2)
SSTO_LASSO = sum((Y.train - mean(Y.train))^2)
n = nrow(data_train)
p = 12
adj_Rsquared_LASSO_trained = 1 - (SSE_LASSO/SSTO_LASSO)*((n-1)/(n-p-1))
adj_Rsquared_LASSO_trained
RMSE_LASSO
```





age and body fat
https://agsjournals.onlinelibrary.wiley.com/doi/abs/10.1111/j.1532-5415.1993.tb06693.x

```{r}
bd_df_adpt = 
  bd_df %>% 
  mutate(age_level = ifelse(age<40,0,ifelse(age<65,1,2)))
  
data_train_adpt<-bd_df_adpt[sub,]
data_test_adpt<-bd_df_adpt[-sub,]

lm_int = lm(bodyfat_brozek ~ age +  age  +weight + neck + abdomen  + thigh  +forearm + wrist+neck:age_level + chest:age_level + abdomen:age_level + hip:age_level + thigh:age_level  +forearm:age_level + wrist:age_level, data = data_train_adpt)

summary(lm_int)

```



```{r}

lm_int_1 = lm(bodyfat_brozek ~  age  +weight + neck + abdomen  + thigh  +forearm + age_level*weight, data = data_train_adpt)
summary(lm_int_1)




lm_int_2 = lm(bodyfat_brozek ~  age  +weight + neck + abdomen  + thigh  +forearm + age_level*neck, data = data_train_adpt)
summary(lm_int_2)



lm_int_3 = lm(bodyfat_brozek ~  age  +weight + neck + abdomen  + thigh  +forearm + age_level*abdomen, data = data_train_adpt)
summary(lm_int_3)



lm_int_4 = lm(bodyfat_brozek ~  age  +weight + neck + abdomen  + thigh  +forearm + age_level*thigh, data = data_train_adpt)
summary(lm_int_4)




lm_int_5 = lm(bodyfat_brozek ~  age  +weight + neck + abdomen  + thigh  +forearm + age_level*forearm, data = data_train_adpt)
summary(lm_int_5)

```



